---
title: "R Notebook"
output: html_notebook
---

This notebook supports as asynchronous lecture meant to introduce you to some basics of spatial data. The lecture will be provided.

Key topics:
-Using here package
-Using the /data-raw/ directory
-Reminder on joins (focus: left join)
Our spatial Data
  -Neighborhoods
  -Joining with non-spatial data
  -Census data
  Joining with spatial data
-Ignore html in git

```{r}
library(knitr)
library(tidyverse)
library(janitor)
library(lubridate)
library(here)
library(sf)
library(tmap)
library(tidycensus)
library(dplyr)
install.packages("gbfs")
library(gbfs)
```


```{r}
neigh=st_read(here("data_raw",
                    "DC_Health_planning_Neighborhoods.geojson")) %>%clean_names()

class(neigh)

```
```{r}
neigh$geometry[1]
neigh$geometry[1]$POLYGON
view(neigh)
plot(neigh)
```

```{r}
df1=tibble(fruit=c("apple", "banana", "cherry"), cost = c(1.5, 1.2,2.25))
df2=tibble(fruit=c("apple", "apple", "cherry", "lemon"),
           dessert=c("pie", "cobbler", "cobbler", "cheesecake"),
           cal=c(400,430,500,550))
df1
```

```{r}
df2
```

```{r}
left_join(df1,df2,by="fruit")
```

##Injesting joining spatial and non-spatial data

```{r message=FALSE, warning=FALSE}
df_c=read_csv(here("data_raw",
              "DC_COVID-19_Total_Positive_Cases_by_Neighborhood.csv")) %>%
  clean_names()

```


```{r}
df_cases=df_c %>%
  filter(as_date(date_reported) =="2021-11-17") %>%
  separate(neighborhood,into=c("code", "name"),sep =":") %>%
  mutate(code=case_when(code=="N35" ~"NO",
                        TRUE ~ code)) 
  

```



```{r}
neigh2=left_join(neigh,df_cases,by=c("code"))

tmap_mode("view")

tm_shape(neigh2) +tm_polygons("total_positives",alpha=.5)
```



```{r eval=FALSE, include=FALSE}
census_api_key("6f39f12de0ff3a29877db82780befcd77e0d5a2a")
```

```{r}
v20 = load_variables(2018, "acs5")
view(v20 )
```

```{r}
df_cencus=get_acs(geography = "tract",
                  variables=c("median_inc"="B06011_001",
                            "pop"="B01001_001",
                            "pop_black"="B02009_001"),
                  state="DC",geometry=TRUE,year=2021) 
                            
```

```{r}
class(df_cencus)
plot(df_cencus)
view(df_cencus)
```

```{r}
df_cens=df_cencus %>%
  
  pivot_wider(names_from = "variable",
              values_from = "estimate")

tm_shape(df_cens) +tm_polygons("median_inc",alpha=.5)
view(df_cens)
```

```{r}
tm_shape(neigh2)+tm_borders(col="blue", lwd=5, alpha=.2)+ tm_shape(df_cens)+tm_borders(col="red", lwd=1, alpha=.3)
```

```{r}
df_j=st_join(df_cens,neigh2)
```
```{r}
df_cens_adj=df_cens %>% st_transform(4326)
```

```{r}
df_j=st_join(df_cens_adj,neigh2,largest=TRUE)
```

```{r}
df_j_rev=st_join(neigh2,df_cens_adj,largest=TRUE)
view(df_j_rev)
```


```{r}
df1=df_j %>% select(median_inc, pop, pop_black,code) %>%
  group_by(code) %>%
  summarize(pop_n=sum(pop),
            pop_black_n=sum(pop_black),
            adj_median_income=sum(pop*median_inc)/pop_n)
plot(df1)
```

```{r}

df2=left_join(neigh2,df1 %>% st_set_geometry(NULL))
```

```{r}
df2=df2 %>% mutate(black_perc=pop_black_n/pop_n, covid_rate=total_positives/pop_n)
tm_shape(df2) +tm_polygons(c("adj_median_income","covid_rate","black_perc"))
view(df2)
```

```{r}
df2 %>% filter(code!="NO") %>%
tm_shape()+tm_polygons(c("adj_median_income","covid_rate","black_perc"),alpha=.4)
```


